---
title: |
  <center> INSTREAM reductions </center>
  <center> Delta F.A.R.M. </center>

author: "Mark Hill"
date: "10/28/2020"
output:
  pdf_document: 
    toc: no
    toc_depth: 3
  html_document: default
---
\newpage
\thispagestyle{empty}
\mbox{}

\tableofcontents

\newpage


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE, message=FALSE}
library(lubridate)
library(tidyverse)
library(dplyr)
library(ggpubr)
library(ggsignif)
library(stats)
library(zoo)
library(scales)
library(gridExtra)
library(grid)
library(mosaic)
library(mosaicData)
library(patchwork)
library(mice)
library(readxl)
```


```{r, echo= FALSE, message= FALSE, warning= FALSE, results= 'hide' }
dat <- read.csv("DF_final.csv",
                stringsAsFactors = FALSE,
                na.strings = c("", "NA") # replaces blanks with NAs
               # skip = 1 # skips to the first row for data
) %>% 
mutate( 
    # changes the formatting of some of the columns to make R happier
    Date = lubridate::mdy(Date),
    # makes the dates as dates R will recognize
    sampleid = as.factor(sampleid),
    # Sets the sampleid column as a factor instead of text strings
    farm = as.factor(farm),
    # Sets the Farm column as a factor instead of text strings
    treatment = as.factor(treatment),
    # Sets the Treatment column as a factor instead of text strings
    TUR_ntu = as.numeric(TUR_ntu),      
    # Sets turbidity as a number instead of text strings 
    #(I have no idea why it wasn't already)
    sample_type = as.factor(sample_type)
  )

# replace underscores in the column headers with periods
names(dat) <- gsub(
  x = names(dat), 
  pattern = "\\_", 
  replacement = "."
)

dat$season <- ifelse(
  month(dat$Date) == 11 | month(dat$Date) ==  12 | month(dat$Date) == 1| 
    month(dat$Date) == 2 |  month(dat$Date) == 3 | month(dat$Date) == 4,
  "cover",
  "cash")

dat <- dat %>% 
  mutate(season = as.factor(season))


# add location data -------------------------------------------------------
 
df_samplesites <- read.csv("df_samplesites.csv")
dat <- left_join(dat, df_samplesites)

# change locations for fields that moved
dat[dat$sampleid == "ARR1",]$lat <-
  ifelse(
    dat[dat$sampleid == "ARR1",]$Date >="2019-10-18",
    33.7475234, # New location 
    33.743923   # old location 
  )
dat[dat$sampleid == "ARR1",]$long <-
  ifelse(
    dat[dat$sampleid == "ARR1",]$Date >="2019-10-18",
    -90.439047, # New location 
    -90.434174   # old location 
  )

dat[dat$sampleid == "ARR2",]$lat <-
  ifelse(
    dat[dat$sampleid == "ARR2",]$Date >="2019-10-18",
    33.747566, # New location 
    33.7439064   # old location 
  )
dat[dat$sampleid == "ARR2",]$long <-
  ifelse(
    dat[dat$sampleid == "ARR2",]$Date >="2019-10-18",
    -90.441001, # New location 
    -90.431302   # old location 
  )

dat[dat$sampleid == "STU1",]$lat <-
  ifelse(
    dat[dat$sampleid == "STU1",]$Date >="2019-12-5",
    33.8177149, # New location 
    33.7921411   # old location 
  )
dat[dat$sampleid == "STU1",]$long <-
  ifelse(
    dat[dat$sampleid == "STU1",]$Date >="2019-12-5",
    -90.252741, # New location 
    -90.267364   # old location 
  )

dat[dat$sampleid == "STU2",]$lat <-
  ifelse(
    dat[dat$sampleid == "STU2",]$Date >="2019-12-5",
    33.8194787, # New location 
    33.7955973   # old location 
  )
dat[dat$sampleid == "STU2",]$long <-
  ifelse(
    dat[dat$sampleid == "STU2",]$Date >="2019-12-5",
    -90.258794, # New location 
    -90.272795   # old location 
  )


specify_decimal <- function(x, k) trimws(format(round(x, k), nsmall=k))
dat$lat <- specify_decimal(dat$lat, 7)
dat$long <- specify_decimal(dat$long, 7)

# LOGIC CHECK -------------------------------------------------------------

# Here the standard columns are checked and any data from dates where the 
# standards didn't meet QAQC objectives are removed.

# First, replace concentration data on bad dates
# list of columns that have concentration data
conc <- c(10,11,13:15,17,19,21,23,25,27,31:39)
  
# This basically says, "For all the columns with concentration data, if the 
# QAQC for the day was bad, set the concentration data to 'NA', if not, then
# do not change it". All of the loops follow the same basic pattern.
for (i in conc) {
  dat[,i] <- ifelse(
    dat[,2] == FALSE,
    NA,
    dat[,i]
  )
}

# Then, replace concentration on individual dates
# list of logical column numbers besides qaqc.check
logic <- c(12, 16, 18, 20, 22, 24, 26)

# Since the check columns are behind the columns they check, we can just change
# the column before the logical one.
for (i in logic) {
  dat[,i - 1] <- ifelse(
    dat[,i] == FALSE,
    NA,
    dat[,i - 1]
  )
}


# replace NO3 values b/c they are tied to NOx check
# we do this seperately b/c it doesn't fit the pattern of the for() loop above
dat[,27] <- ifelse(
  dat[,24] == FALSE,
  NA,
  dat[,27]
)

# Next we replace data in the loads columns. Since they are spaced differently,
# it has to be done in 3 different loops.

# replace tss loads
for (i in c(21, 24, 27)) { 
  dat[,10 + i] <- ifelse(
    is.na(dat[,10]) == TRUE,
    NA,
    dat[,10 + i]
  )
}

# replace tn loads
for (i in c(17, 20, 23)) { 
  dat[,15 + i] <- ifelse(
    is.na(dat[,15]) == TRUE,
    NA,
    dat[,15 + i]
  )
}

# replace tip loads
for (i in c(16, 19, 22)) { 
  dat[,17 + i] <- ifelse(
    is.na(dat[,17]) == TRUE,
    NA,
    dat[,17 + i]
  )
}

# Finally, we remove the columns that have the logical checks (because I
# the rest of the code relies on absolute places in the dataframe).
# remove logical columns
dat <- dat %>% 
  select(-c(2, 12, 16, 18, 20, 22, 24, 26))

rm(conc, i, logic)


# Impute missing values ---------------------------------------------------
#dat <- dat %>%
#  mutate(
#    farm = as.character(farm),
#    treatment = as.character(treatment)
#  )
#
#dat$treatment <-
#  ifelse(
#    is.na(dat$treatment),
#    "INST",
#    dat$treatment
#  )
#dat$farm <-
#  ifelse(
#    is.na(dat$farm),
#    "INST",
#    dat$farm
#  )
#
#dat <- mutate(
#  dat,
#  farm = as.factor(farm),
#  treatment = as.factor(treatment)
#)
#
# ..Sediment ---------------------------

impdat <-
  mice(
    data = dat[,c(9, 10, 14)],
    m = 5,
    maxit = 50,
    method = 'pmm',
    seed = 1234,
    printFlag = FALSE
  )

set.seed(1234)  # set seed for random imputed dataset selection
compdat <- complete(impdat, round(runif(1, 1, 5), 0))
# randomly select imputed dataset and combine with complete dataset

# inspect imputed data
# imputed datasets in magenta; observed in blue. are values plausible?
#stripplot(impdat)
# do imputed points fall in likely area of distrib.?

dat[,c(9, 10, 14)] <- compdat[,1:ncol(compdat)]

# ..Ortho P -------------------------


impdat <-
  mice(
    data = dat[,c(14, 16)],
    m = 5,
    maxit = 50,
    method = 'pmm',
    seed = 1234,
    printFlag = FALSE
  )

set.seed(1234)  # set seed for random imputed dataset selection
compdat <- complete(impdat, round(runif(1, 1, 5), 0))
# randomly select imputed dataset and combine with complete dataset

# inspect imputed data
# imputed datasets in magenta; observed in blue. are values plausible?
# stripplot(impdat)
# do imputed points fall in likely area of distrib.?

dat[,c(14, 16)] <- compdat[,1:ncol(compdat)]

# ..Nitrogen ------------------------------------

impdat <-
  mice(
    data = dat[,c(11:13, 15, 17)],
    m = 5,
    maxit = 50,
    method = 'pmm',
    seed = 1234,
    printFlag = FALSE
  )

set.seed(1234)  # set seed for random imputed dataset selection
compdat <- complete(impdat, round(runif(1, 1, 5), 0))
# randomly select imputed dataset and combine with complete dataset

# inspect imputed data
# imputed datasets in magenta; observed in blue. are values plausible?
#stripplot(impdat)
# do imputed points fall in likely area of distrib.?

dat[,c(11:13, 15, 17)] <- compdat[,1:ncol(compdat)]

# ..No3 -----------------------------------

impdat <-
  mice(
    data = dat[,c(11:13, 15, 19)],
    m = 5,
    maxit = 50,
    method = 'pmm',
    seed = 1234,
    printFlag = FALSE
  )

set.seed(1234)  # set seed for random imputed dataset selection
compdat <- complete(impdat, round(runif(1, 1, 5), 0))
# randomly select imputed dataset and combine with complete dataset

# inspect imputed data
# imputed datasets in magenta; observed in blue. are values plausible?
#stripplot(impdat)
# do imputed points fall in likely area of distrib.?

dat[,c(11:13, 15, 19)] <- compdat[,1:ncol(compdat)]

rm(impdat)
rm(compdat)

# repair MDLs -------------------------------------------------------------
if (sum(dat$TSS.mgl < 5, na.rm = TRUE) / sum(!is.na(dat$TSS.mgl)) < 0.05) {
  print("Less than 5% of TSS values are below MDL")
  print(sum(dat$TSS.mgl < 5, na.rm = TRUE) / sum(!is.na(dat$TSS.mgl)) *100)
  dat <- dat %>% mutate(TSS.mgl = replace(TSS.mgl, TSS.mgl < 5, 5/2))
} else {
  print("More than 5% of TSS values are below MDL and")
  print("the values have remained unchanged")
  print(dat %>% count(TSS.mgl < 5))
  print(sum(dat$TSS.mgl < 5, na.rm = TRUE) / sum(!is.na(dat$TSS.mgl)) *100)}


if (sum(dat$TUR.ntu < .02, na.rm = TRUE) / sum(!is.na(dat$TUR.ntu)) < 0.05) {
  print("Less than 5% of TUR values are below MDL")
  print(sum(dat$TUR.ntu < .02, na.rm = TRUE) / sum(!is.na(dat$TUR.ntu)) *100)
  dat <- dat %>% mutate(TUR.ntu = replace(TUR.ntu, TUR.ntu < .02, .02/2))
} else {
  print("More than 5% of TUR values are below MDL and")
  print("the values have remained unchanged")
  print(dat %>% count(TUR.ntu < .02))
  print(sum(dat$TUR.ntu < .02, na.rm = TRUE) / sum(!is.na(dat$TUR.ntu)) *100)}


if (sum(dat$NO3.NO2.mgl < .23, na.rm = TRUE) / sum(!is.na(dat$NO3.NO2.mgl)) < 0.05) {
  print("Less than 5% of NO3.NO2 values are below MDL")
  print(sum(dat$NO3.NO2.mgl < .23, na.rm = TRUE) / sum(!is.na(dat$NO3.NO2.mgl)) *100)
  dat <- dat %>% mutate(NO3.NO2.mgl = replace(NO3.NO2.mgl, NO3.NO2.mgl < .23, .23/2))
} else {
  print("More than 5% of NO3.NO2 values are below MDL and")
  print("the values have remained unchanged")
  print(dat %>% count(NO3.NO2.mgl < .23))
  print(sum(dat$NO3.NO2.mgl < .23, na.rm = TRUE) / sum(!is.na(dat$NO3.NO2.mgl)) *100)}


if (sum(dat$TKN.mgl < .77, na.rm = TRUE) / sum(!is.na(dat$TKN.mgl)) < 0.05) {
  print("Less than 5% of TKN values are below MDL")
  print(sum(dat$TKN.mgl < .77, na.rm = TRUE) / sum(!is.na(dat$TKN.mgl)) *100)
  dat <- dat %>% mutate(TKN.mgl = replace(TKN.mgl, TKN.mgl < .77, .77/2))
} else {
  print("More than 5% of TKN values are below MDL and")
  print("the values have remained unchanged")
  print(dat %>% count(TKN.mgl < .77))
  print(sum(dat$TKN.mgl < .77, na.rm = TRUE) / sum(!is.na(dat$TKN.mgl)) * 100)
  }


if (sum(dat$TN.mgl < 1, na.rm = TRUE) / sum(!is.na(dat$TN.mgl)) < 0.05) {
  print("Less than 5% of TN values are below MDL")
  print(sum(dat$TN.mgl < 1, na.rm = TRUE) / sum(!is.na(dat$TN.mgl)) *100)
  dat <- dat %>% mutate(TN.mgl = replace(TN.mgl, TN.mgl < 1, 1/2))
} else {
  print("More than 5% of TN values are below MDL and")
  print("the values have remained unchanged")
  print(dat %>% count(TN.mgl < 1))
  print(sum(dat$TN.mgl < 1, na.rm = TRUE) / sum(!is.na(dat$TN.mgl)) *100)
  }


if (sum(dat$TIP.mgl < .05, na.rm = TRUE) / sum(!is.na(dat$TIP.mgl)) < 0.05) {
  print("Less than 5% of TIP values are below MDL")
  print(sum(dat$TIP.mgl < .05, na.rm = TRUE) / sum(!is.na(dat$TIP.mgl)) *100)
  dat <- dat %>% mutate(TIP.mgl = replace(TIP.mgl, TIP.mgl < .05, .05/2))
} else {
  print("More than 5% of TIP values are below MDL and")
  print("the values have remained unchanged")
  print(dat %>% count(TIP.mgl < .05))
  print(sum(dat$TIP.mgl < .05, na.rm = TRUE) / sum(!is.na(dat$TIP.mgl)) *100)}


if (sum(dat$NH3.mgl < .01, na.rm = TRUE) / sum(!is.na(dat$NH3.mgl)) < 0.05) {
  print("Less than 5% of NH3 values are below MDL")
  print(sum(dat$NH3.mgl < .01, na.rm = TRUE) / sum(!is.na(dat$NH3.mgl)) *100)
  dat <- dat %>% mutate(NH3.mgl = replace(NH3.mgl, NH3.mgl < .01, .01/2))
} else {
  print("More than 5% of NH3 values are below MDL and")
  print("the values have remained unchanged")
  print(dat %>% count(NH3.mgl < .01))
  print(sum(dat$NH3.mgl < .01, na.rm = TRUE) / sum(!is.na(dat$NH3.mgl)) *100)}


if (sum(dat$OrthoP.mgl < .01, na.rm = TRUE) / sum(!is.na(dat$OrthoP.mgl)) < 0.05) {
  print("Less than 5% of Ortho Phosphorus values are below MDL")
  print(sum(dat$OrthoP.mgl < .01, na.rm = TRUE) / sum(!is.na(dat$OrthoP.mgl)) *100)
  dat <- dat %>% mutate(OrthoP.mgl = replace(OrthoP.mgl, OrthoP.mgl < .01, .01/2))
} else {
  print("More than 5% of Ortho Phosphorus values are below MDL and")
  print("the values have remained unchanged")
  print(dat %>% count(OrthoP.mgl < .01))
  print(sum(dat$OrthoP.mgl < .01, na.rm = TRUE) / sum(!is.na(dat$OrthoP.mgl)) *100)}


if (sum(dat$NOx.mgl < .007, na.rm = TRUE) / sum(!is.na(dat$NOx.mgl)) < 0.05) {
  print("Less than 5% of NOx values are below MDL")
  print(sum(dat$NOx.mgl < .007, na.rm = TRUE) / sum(!is.na(dat$NOx.mgl)) *100)
  dat <- dat %>% mutate(NOx.mgl = replace(NOx.mgl, NOx.mgl < .007, .007/2))
} else {
  print("More than 5% of NOx values are below MDL and")
  print("the values have remained unchanged")
  print(dat %>% count(NOx.mgl < .007))
  print(sum(dat$NOx.mgl < .007, na.rm = TRUE) / sum(!is.na(dat$NOx.mgl)) *100)}


if (sum(dat$NO2.mgl < .05, na.rm = TRUE) / sum(!is.na(dat$NO2.mgl)) < 0.05) {
  print("Less than 5% of NO2 values are below MDL")
  print(sum(dat$NO2.mgl < .05, na.rm = TRUE) / sum(!is.na(dat$NO2.mgl)) *100)
  dat <- dat %>% mutate(NO2.mgl = replace(NO2.mgl, NO2.mgl < .05, .05/2))
} else {
  print("More than 5% of NO2 values are below MDL and")
  print("the values have remained unchanged")
  print(dat %>% count(NO2.mgl < .05))
  print(sum(dat$NO2.mgl < .05, na.rm = TRUE) / sum(!is.na(dat$NO2.mgl)) *100)}


if (sum(dat$NO3.mgl < .0007, na.rm = TRUE) / sum(!is.na(dat$NO3.mgl)) < 0.05) {
  print("Less than 5% of NO3 values are below MDL")
  print(sum(dat$NO3.mgl < .0007, na.rm = TRUE) / sum(!is.na(dat$NO3.mgl)) *100)
  dat <- dat %>% mutate(NO3.mgl = replace(NO3.mgl, NO3.mgl < .0007, .0007/2))
} else {
  print("More than 5% of NO3 values are below MDL and")
  print("the values have remained unchanged")
  print(dat %>% count(NO3.mgl < .0007))
  print(sum(dat$NO3.mgl < .0007, na.rm = TRUE) / sum(!is.na(dat$NO3.mgl)) *100)}

#another round of imput for ammonia and tkn
#NH3 ---------------

dat$NH3.mgl <-
  ifelse(
    dat$NH3.mgl < .01,
    NA,
    dat$NH3.mgl
  )

impdat <-
  mice(
    data = dat[,c(11:13, 15, 17)],
    m = 5,
    maxit = 50,
    method = 'pmm',
    seed = 1234,
    printFlag = FALSE
  )

set.seed(1234)  # set seed for random imputed dataset selection
compdat <- complete(impdat, round(runif(1, 1, 5), 0))
# randomly select imputed dataset and combine with complete dataset

# inspect imputed data
# imputed datasets in magenta; observed in blue. are values plausible?
#stripplot(impdat)
# do imputed points fall in likely area of distrib.?

dat[,c(11:13, 15, 17)] <- compdat[,1:ncol(compdat)]

# TKN-------------------
print(
  paste(
    sum(dat$TKN.mgl < 0, na.rm = TRUE) / sum(!is.na(dat$TKN.mgl)) *100,
    "% of TKN values are below 0"
  )
)


dat$TKN.mgl <-
  ifelse(
    dat$TKN.mgl < 0,
    NA,
    dat$TKN.mgl
  )

summary(dat$TKN.mgl)

impdat <-
  mice(
    data = dat[,c(11:13, 15, 17)],
    m = 5,
    maxit = 50,
    method = 'pmm',
    seed = 1234,
    printFlag = FALSE
  )

set.seed(1234)  # set seed for random imputed dataset selection
compdat <- complete(impdat, round(runif(1, 1, 5), 0))
# randomly select imputed dataset and combine with complete dataset

# inspect imputed data
# imputed datasets in magenta; observed in blue. are values plausible?
#stripplot(impdat)
# do imputed points fall in likely area of distrib.?

dat[,c(11:13, 15, 17)] <- compdat[,1:ncol(compdat)]



```

```{r, echo= FALSE, message= FALSE, warning= FALSE}
INST <- dat[dat$sample.type== "INST",]%>% 
  rename(LOCCODE = sampleid,
         TSS= TSS.mgl,
         TUR = TUR.ntu,
         NN = NO3.NO2.mgl,
         TKN= TKN.mgl,
         TN = TN.mgl,
         TIP = TIP.mgl,
         NH3 = NH3.mgl,
         DIP = OrthoP.mgl
         )

oldINST <- read.csv("combinedgrabs.csv") %>% 
 rename(Date = 1,
        TUR = 13)  %>% 
  mutate(
         Date = lubridate::mdy(Date),
         Alk = as.numeric(Alk),
         NH3 = as.numeric(NH3),
         TSS = as.numeric(TSS),
         NN = as.numeric(NN),
         TP = as.numeric(TP),
         TN = as.numeric(TN),
         OrgN = as.numeric(OrgN),
         TOC = as.numeric(TOC),
         COD = as.numeric(COD),
         Cl = as.numeric(Cl),
         TKN = as.numeric(TKN)
         )
  
test <- full_join(INST, oldINST)
test <- test[,-c(4:8, 20:31, 39:41)]
test$LOCCODE <- as.factor(test$LOCCODE)


attach(test)
oldharris <- test[LOCCODE== "HN3"|LOCCODE== "HN4"|LOCCODE== "HN5"
                 |LOCCODE== "HN6"|LOCCODE== "HNO"|LOCCODE== "HNSTA"
                 |LOCCODE== "HNT"|LOCCODE== "HS1"|LOCCODE== "HS2",]

oldporter <- test[LOCCODE== "PN1"|LOCCODE== "PN2"|LOCCODE== "PS1"|
                  LOCCODE== "PS2"|LOCCODE== "PS3"|LOCCODE== "PS4"|
                  LOCCODE== "PS5"|LOCCODE== "PS6",]
detach(test)

```

\newpage

# TURBIDITY

## Harris Bayou

### Historical Harris Bayou sites

```{r, warning= FALSE, echo=FALSE}
summarise(oldharris, n = count(!is.na(oldharris$TUR)), Mean= mean(TUR, na.rm=TRUE), Median= median(TUR, na.rm = TRUE))
```

### Data from Harris Bayou site associated with the original Delta FARM project

```{r, echo=FALSE}
test %>% 
  filter(LOCCODE =="HB") %>% 
summarise(n= count(!is.na(TUR)),
          Mean= mean(TUR, na.rm= TRUE),
          Median= median(TUR, na.rm = TRUE))
```

### Plot the data points. Red is our recent data.

```{r fig1, results='hold', warning=FALSE, echo= FALSE, fig.height= 3 }
ggplot()+
  geom_jitter(data= oldharris, aes(y= TUR, x= LOCCODE))+
  geom_jitter(data= test[test$LOCCODE == "HB",], aes(y= TUR, x= LOCCODE, color= "red"))+
  theme(legend.position = "none")
  
```

```{r fig2, warning=FALSE, fig.height= 3, echo= FALSE}
a <- ggplot()+
  geom_boxplot(data= oldharris, aes(y= TUR, x= LOCCODE))+
  geom_boxplot(data= test[test$LOCCODE == "HB",], aes(y= TUR, x= LOCCODE, color= "red"))+
  theme(legend.position = "none", axis.text.x = element_text(angle = 45))
b <- ggplot()+
  geom_boxplot(data= test[test$LOCCODE == "HB",], aes(y= TUR, x= LOCCODE, color= "red"))+
  geom_boxplot(data= oldharris, aes(y= TUR))+
  theme(legend.position = "none", axis.title.y = element_blank())

a+b
```

\newpage

### Test for a difference in means

```{r}
wilcox.test(y= oldharris$TUR, x= test[test$LOCCODE == "HB", "TUR"], alternative = "less")
```

> Small P-value here indicates that we can reject the null hypothesis that the difference is not less than zero and conclude, at a %95 confidence level, that the values of our recent data is in fact lower than the historical data.

### Percent reduction in the median value
```{r}
x <- median(oldharris$TUR, na.rm= TRUE)
y <- median(test[test$LOCCODE == "HB", "TUR"])
(y-x)/x * 100

```
 
 > Here we would use median values because there are outliers and apparent skewness from the boxplot. The mean reduction is even greater (%72).

\newpage

# TURBIDITY

## Porter Bayou

### Historical Porter Bayou sites

```{r, warning= FALSE, echo=FALSE}
summarise(oldporter, n = count(!is.na(oldporter$TUR)), Mean= mean(TUR, na.rm=TRUE), Median= median(TUR, na.rm = TRUE))
```

### Data from Porter Bayou site associated with the original Delta FARM project

```{r, echo=FALSE}
test %>% 
  filter(LOCCODE =="PB") %>% 
summarise(n= count(!is.na(TUR)),
          Mean= mean(TUR, na.rm= TRUE), 
          Median= median(TUR, na.rm = TRUE))
```

### Plot the data points. Red is our recent data.

```{r fig3, results='hold', warning=FALSE, echo= FALSE, fig.height= 3}
ggplot()+
  geom_jitter(data= oldporter, aes(y= TUR, x= LOCCODE))+
  geom_jitter(data= test[test$LOCCODE == "PB",], aes(y= TUR, x= LOCCODE, color= "red"))+
  theme(legend.position = "none")
  
```

```{r fig4, warning=FALSE, fig.height=3, echo= FALSE}
a <- ggplot()+
  geom_boxplot(data= oldporter, aes(y= TUR, x= LOCCODE))+
  geom_boxplot(data= test[test$LOCCODE == "PB",], aes(y= TUR, x= LOCCODE, color= "red"))+
  theme(legend.position = "none", axis.text.x = element_text(angle = 45))
b <- ggplot()+
  geom_boxplot(data= test[test$LOCCODE == "PB",], aes(y= TUR, x= LOCCODE, color= "red"))+
  geom_boxplot(data= oldporter, aes(y= TUR))+
  theme(legend.position = "none", axis.title.y = element_blank())

a+b
```

\newpage

### Test for a difference in means

```{r}
wilcox.test(y= oldporter$TUR, x= test[test$LOCCODE == "PB", "TUR"], alternative = "greater")
```

> Small P-value here indicates that we can reject the null hypothesis that the difference is not greater than zero and conclude, at a %95 confidence level, that the values of our recent data is in fact higher than the historical data.

### Percent reduction in the median value

```{r}
x <- median(oldporter$TUR, na.rm= TRUE)
y <- median(test[test$LOCCODE == "PB", "TUR"])
(y-x)/x * 100

```
 
 > Here we would use median values because there are outliers and apparent skewness from the boxplot. Postive values represents increase.

\newpage

# TOTAL SUSPENDED SOLIDS

## Harris Bayou

### Historical Harris Bayou sites

```{r, warning= FALSE, echo=FALSE}
summarise(oldharris, n = count(!is.na(oldharris$TSS)), Mean= mean(TSS, na.rm=TRUE), Median= median(TSS, na.rm = TRUE))
```

### Data from Harris Bayou site associated with the original Delta FARM project

```{r, echo=FALSE}
test %>% 
  filter(LOCCODE =="HB") %>% 
summarise(n= count(!is.na(TSS)),
          Mean= mean(TSS, na.rm= TRUE),
          Median= median(TSS, na.rm = TRUE))
```

### Plot the data points. Red is our recent data.

```{r fig5, results='hold', warning=FALSE, echo= FALSE, fig.height= 3 }
ggplot()+
  geom_jitter(data= oldharris, aes(y= TSS, x= LOCCODE))+
  geom_jitter(data= test[test$LOCCODE == "HB",], aes(y= TSS, x= LOCCODE, color= "red"))+
  theme(legend.position = "none")
  
```

```{r fig6, warning=FALSE, fig.height= 3, echo= FALSE}
a <- ggplot()+
  geom_boxplot(data= oldharris, aes(y= TSS, x= LOCCODE))+
  geom_boxplot(data= test[test$LOCCODE == "HB",], aes(y= TSS, x= LOCCODE, color= "red"))+
  theme(legend.position = "none", axis.text.x = element_text(angle = 45))
b <- ggplot()+
  geom_boxplot(data= test[test$LOCCODE == "HB",], aes(y= TSS, x= LOCCODE, color= "red"))+
  geom_boxplot(data= oldharris, aes(y= TSS))+
  theme(legend.position = "none", axis.title.y = element_blank())

a+b
```

\newpage

### Test for a difference in means

```{r}
wilcox.test(y= oldharris$TSS, x= test[test$LOCCODE == "HB", "TSS"], alternative = "less")
```

> P value of less than alpha = 0.05 means we can conclude the alternative hypothesis. 

### Percent reduction in the median value
```{r}
x <- median(oldharris$TSS, na.rm= TRUE)
y <- median(test[test$LOCCODE == "HB", "TSS"])
(y-x)/x * 100

```
 
 > 

\newpage

# TOTAL SUSPENDED SOLIDS

## Porter Bayou

### Historical Porter Bayou sites

```{r, warning= FALSE, echo=FALSE}
summarise(oldporter, n = count(!is.na(oldporter$TSS)), Mean= mean(TSS, na.rm=TRUE), Median= median(TSS, na.rm = TRUE))
```

### Data from Porter Bayou site associated with the original Delta FARM project

```{r, echo=FALSE}
test %>% 
  filter(LOCCODE =="PB") %>% 
summarise(n= count(!is.na(TSS)),
          Mean= mean(TSS, na.rm= TRUE), 
          Median= median(TSS, na.rm = TRUE))
```

### Plot the data points. Red is our recent data.

```{r fig7, results='hold', warning=FALSE, echo= FALSE, fig.height= 3}
ggplot()+
  geom_jitter(data= oldporter, aes(y= TSS, x= LOCCODE))+
  geom_jitter(data= test[test$LOCCODE == "PB",], aes(y= TSS, x= LOCCODE, color= "red"))+
  theme(legend.position = "none")+
  coord_cartesian(ylim = c(0,2500))
```

```{r fig8, warning=FALSE, fig.height=3, echo= FALSE}
a <- ggplot()+
  geom_boxplot(data= oldporter, aes(y= TSS, x= LOCCODE))+
  geom_boxplot(data= test[test$LOCCODE == "PB",], aes(y= TSS, x= LOCCODE, color= "red"))+
  theme(legend.position = "none", axis.text.x = element_text(angle = 45))+
    coord_cartesian(ylim = c(0,2500))

b <- ggplot()+
  geom_boxplot(data= test[test$LOCCODE == "PB",], aes(y= TSS, x= LOCCODE, color= "red"))+
  geom_boxplot(data= oldporter, aes(y= TSS))+
  theme(legend.position = "none", axis.title.y = element_blank())+
    coord_cartesian(ylim = c(0,2500))


a+b
```
  * some outliers clipped from visualization
  
\newpage

### Test for a difference in means

```{r}
wilcox.test(y= oldporter$TSS, x= test[test$LOCCODE == "PB", "TSS"], alternative = "greater")
```

> The test lends statistical significance to old porter bayou data values being less than what we recently collected.

### Percent reduction in the median value

```{r}
x <- median(oldporter$TSS, na.rm= TRUE)
y <- median(test[test$LOCCODE == "PB", "TSS"])
(y-x)/x * 100

```
 
 >This positive number represents a four fold increase from values formally observed in porter bayou.

\newpage

# TOTAL NITROGEN

## Harris Bayou

### Historical Harris Bayou sites

```{r, warning= FALSE, echo=FALSE}
summarise(oldharris, n = count(!is.na(oldharris$TN)), Mean= mean(TN, na.rm=TRUE), Median= median(TN, na.rm = TRUE))
```

### Data from Harris Bayou site associated with the original Delta FARM project

```{r, echo=FALSE}
test %>% 
  filter(LOCCODE =="HB") %>% 
summarise(n= count(!is.na(TN)),
          Mean= mean(TN, na.rm= TRUE),
          Median= median(TN, na.rm = TRUE))
```

### Plot the data points. Red is our recent data.

```{r fig9, results='hold', warning=FALSE, echo= FALSE, fig.height= 3 }
ggplot()+
  geom_jitter(data= oldharris, aes(y= TN, x= LOCCODE))+
  geom_jitter(data= test[test$LOCCODE == "HB",], aes(y= TN, x= LOCCODE, color= "red"))+
  theme(legend.position = "none")
  
```

```{r fig10, warning=FALSE, fig.height= 3, echo= FALSE}
a <- ggplot()+
  geom_boxplot(data= oldharris, aes(y= TN, x= LOCCODE))+
  geom_boxplot(data= test[test$LOCCODE == "HB",], aes(y= TN, x= LOCCODE, color= "red"))+
  theme(legend.position = "none", axis.text.x = element_text(angle = 45))
b <- ggplot()+
  geom_boxplot(data= test[test$LOCCODE == "HB",], aes(y= TN, x= LOCCODE, color= "red"))+
  geom_boxplot(data= oldharris, aes(y= TN))+
  theme(legend.position = "none", axis.title.y = element_blank())

a+b
```

\newpage

### Test for a difference in means

```{r}
wilcox.test(y= oldharris$TN, x= test[test$LOCCODE == "HB", "TN"], alternative = "less")
```

>  Pvalue < alpha of 0.5: reject null that old values are <= new observations.                    Ha: new observations < old oberservations.

### Percent reduction in the median value
```{r}
x <- median(oldharris$TN, na.rm= TRUE)
y <- median(test[test$LOCCODE == "HB", "TN"])
(y-x)/x * 100

```
 
 >  negative value represents a twenty percent reduction over the historical observations

\newpage

# TOTAL NITROGEN

## Porter Bayou

### Historical Porter Bayou sites

```{r, warning= FALSE, echo=FALSE}
summarise(oldporter, n = count(!is.na(oldporter$TN)), Mean= mean(TN, na.rm=TRUE), Median= median(TN, na.rm = TRUE))
```

### Data from Porter Bayou site associated with the original Delta FARM project

```{r, echo=FALSE}
test %>% 
  filter(LOCCODE =="PB") %>% 
summarise(n= count(!is.na(TN)),
          Mean= mean(TN, na.rm= TRUE), 
          Median= median(TN, na.rm = TRUE))
```

### Plot the data points. Red is our recent data.

```{r fig11, results='hold', warning=FALSE, echo= FALSE, fig.height= 3}
ggplot()+
  geom_jitter(data= oldporter, aes(y= TN, x= LOCCODE))+
  geom_jitter(data= test[test$LOCCODE == "PB",], aes(y= TN, x= LOCCODE, color= "red"))+
  theme(legend.position = "none")
  #coord_cartesian(ylim = c(0,2500))
```

```{r fig12, warning=FALSE, fig.height=3, echo= FALSE}
a <- ggplot()+
  geom_boxplot(data= oldporter, aes(y= TN, x= LOCCODE))+
  geom_boxplot(data= test[test$LOCCODE == "PB",], aes(y= TN, x= LOCCODE, color= "red"))+
  theme(legend.position = "none", axis.text.x = element_text(angle = 45))
    #coord_cartesian(ylim = c(0,2500))

b <- ggplot()+
  geom_boxplot(data= test[test$LOCCODE == "PB",], aes(y= TN, x= LOCCODE, color= "red"))+
  geom_boxplot(data= oldporter, aes(y= TN))+
  theme(legend.position = "none", axis.title.y = element_blank())
  #coord_cartesian(ylim = c(0,2500))


a+b
```

\newpage

### Test for a difference in means

```{r}
wilcox.test(y= oldporter$TN, x= test[test$LOCCODE == "PB", "TN"], alternative = "greater")
```

> Reject null that porter bayou TN concentrations are <= old observations. <b> Ha: new observations are greater than old ones.

### Percent reduction in the median value

```{r}
x <- median(oldporter$TN, na.rm= TRUE)
y <- median(test[test$LOCCODE == "PB", "TN"])
(y-x)/x * 100

```
 
 > postive values represent increases






