---
title: |
  
  <center> Farmer to Farmer EPA Gulf of Mexico Program </center>
  <center> and </center>
  <center> Delta F.A.R.M. Project </center>
  <center> *Cropping Systems for Improving Farm Profitability and Water Quality* </center>
  <center> INSTREAM reductions </center>


author: "Produced in coordination with Delta F.A.R.M. and Mississippi State University"
date: "11/5/2020"
documentclass: article

geometry: margin=2cm
output:
  pdf_document: 
    toc: true
    toc_depth: 2

  html_document: default
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE, message=FALSE}
library(lubridate)
library(tidyverse)
library(dplyr)
library(ggpubr)
library(ggsignif)
library(stats)
library(zoo)
library(scales)
library(gridExtra)
library(grid)
library(mosaic)
library(mosaicData)
library(patchwork)
library(mice)
library(readxl)
library(kableExtra)
```


```{r, echo= FALSE, message= FALSE, warning= FALSE, results= 'hide' }
dat <- read.csv("DF_final.csv",
                stringsAsFactors = FALSE,
                na.strings = c("", "NA") # replaces blanks with NAs
               # skip = 1 # skips to the first row for data
) %>% 
mutate( 
    # changes the formatting of some of the columns to make R happier
    Date = lubridate::mdy(Date),
    # makes the dates as dates R will recognize
    sampleid = as.factor(sampleid),
    # Sets the sampleid column as a factor instead of text strings
    farm = as.factor(farm),
    # Sets the Farm column as a factor instead of text strings
    treatment = as.factor(treatment),
    # Sets the Treatment column as a factor instead of text strings
    TUR_ntu = as.numeric(TUR_ntu),      
    # Sets turbidity as a number instead of text strings 
    #(I have no idea why it wasn't already)
    sample_type = as.factor(sample_type)
  )

# replace underscores in the column headers with periods
names(dat) <- gsub(
  x = names(dat), 
  pattern = "\\_", 
  replacement = "."
)

dat$season <- ifelse(
  month(dat$Date) == 11 | month(dat$Date) ==  12 | month(dat$Date) == 1| 
    month(dat$Date) == 2 |  month(dat$Date) == 3 | month(dat$Date) == 4,
  "cover",
  "cash")

dat <- dat %>% 
  mutate(season = as.factor(season))


# add location data -------------------------------------------------------
 
df_samplesites <- read.csv("df_samplesites.csv")
dat <- left_join(dat, df_samplesites)

# change locations for fields that moved
dat[dat$sampleid == "ARR1",]$lat <-
  ifelse(
    dat[dat$sampleid == "ARR1",]$Date >="2019-10-18",
    33.7475234, # New location 
    33.743923   # old location 
  )
dat[dat$sampleid == "ARR1",]$long <-
  ifelse(
    dat[dat$sampleid == "ARR1",]$Date >="2019-10-18",
    -90.439047, # New location 
    -90.434174   # old location 
  )

dat[dat$sampleid == "ARR2",]$lat <-
  ifelse(
    dat[dat$sampleid == "ARR2",]$Date >="2019-10-18",
    33.747566, # New location 
    33.7439064   # old location 
  )
dat[dat$sampleid == "ARR2",]$long <-
  ifelse(
    dat[dat$sampleid == "ARR2",]$Date >="2019-10-18",
    -90.441001, # New location 
    -90.431302   # old location 
  )

dat[dat$sampleid == "STU1",]$lat <-
  ifelse(
    dat[dat$sampleid == "STU1",]$Date >="2019-12-5",
    33.8177149, # New location 
    33.7921411   # old location 
  )
dat[dat$sampleid == "STU1",]$long <-
  ifelse(
    dat[dat$sampleid == "STU1",]$Date >="2019-12-5",
    -90.252741, # New location 
    -90.267364   # old location 
  )

dat[dat$sampleid == "STU2",]$lat <-
  ifelse(
    dat[dat$sampleid == "STU2",]$Date >="2019-12-5",
    33.8194787, # New location 
    33.7955973   # old location 
  )
dat[dat$sampleid == "STU2",]$long <-
  ifelse(
    dat[dat$sampleid == "STU2",]$Date >="2019-12-5",
    -90.258794, # New location 
    -90.272795   # old location 
  )


specify_decimal <- function(x, k) trimws(format(round(x, k), nsmall=k))
dat$lat <- specify_decimal(dat$lat, 7)
dat$long <- specify_decimal(dat$long, 7)

# LOGIC CHECK -------------------------------------------------------------

# Here the standard columns are checked and any data from dates where the 
# standards didn't meet QAQC objectives are removed.

# First, replace concentration data on bad dates
# list of columns that have concentration data
conc <- c(10,11,13:15,17,19,21,23,25,27,31:39)
  
# This basically says, "For all the columns with concentration data, if the 
# QAQC for the day was bad, set the concentration data to 'NA', if not, then
# do not change it". All of the loops follow the same basic pattern.
for (i in conc) {
  dat[,i] <- ifelse(
    dat[,2] == FALSE,
    NA,
    dat[,i]
  )
}

# Then, replace concentration on individual dates
# list of logical column numbers besides qaqc.check
logic <- c(12, 16, 18, 20, 22, 24, 26)

# Since the check columns are behind the columns they check, we can just change
# the column before the logical one.
for (i in logic) {
  dat[,i - 1] <- ifelse(
    dat[,i] == FALSE,
    NA,
    dat[,i - 1]
  )
}


# replace NO3 values b/c they are tied to NOx check
# we do this seperately b/c it doesn't fit the pattern of the for() loop above
dat[,27] <- ifelse(
  dat[,24] == FALSE,
  NA,
  dat[,27]
)

# Next we replace data in the loads columns. Since they are spaced differently,
# it has to be done in 3 different loops.

# replace tss loads
for (i in c(21, 24, 27)) { 
  dat[,10 + i] <- ifelse(
    is.na(dat[,10]) == TRUE,
    NA,
    dat[,10 + i]
  )
}

# replace tn loads
for (i in c(17, 20, 23)) { 
  dat[,15 + i] <- ifelse(
    is.na(dat[,15]) == TRUE,
    NA,
    dat[,15 + i]
  )
}

# replace tip loads
for (i in c(16, 19, 22)) { 
  dat[,17 + i] <- ifelse(
    is.na(dat[,17]) == TRUE,
    NA,
    dat[,17 + i]
  )
}

# Finally, we remove the columns that have the logical checks (because I
# the rest of the code relies on absolute places in the dataframe).
# remove logical columns
dat <- dat %>% 
  select(-c(2, 12, 16, 18, 20, 22, 24, 26))

rm(conc, i, logic)


# Impute missing values ---------------------------------------------------
#dat <- dat %>%
#  mutate(
#    farm = as.character(farm),
#    treatment = as.character(treatment)
#  )
#
#dat$treatment <-
#  ifelse(
#    is.na(dat$treatment),
#    "INST",
#    dat$treatment
#  )
#dat$farm <-
#  ifelse(
#    is.na(dat$farm),
#    "INST",
#    dat$farm
#  )
#
#dat <- mutate(
#  dat,
#  farm = as.factor(farm),
#  treatment = as.factor(treatment)
#)
#
# ..Sediment ---------------------------

impdat <-
  mice(
    data = dat[,c(9, 10, 14)],
    m = 5,
    maxit = 50,
    method = 'pmm',
    seed = 1234,
    printFlag = FALSE
  )

set.seed(1234)  # set seed for random imputed dataset selection
compdat <- complete(impdat, round(runif(1, 1, 5), 0))
# randomly select imputed dataset and combine with complete dataset

# inspect imputed data
# imputed datasets in magenta; observed in blue. are values plausible?
#stripplot(impdat)
# do imputed points fall in likely area of distrib.?

dat[,c(9, 10, 14)] <- compdat[,1:ncol(compdat)]

# ..Ortho P -------------------------


impdat <-
  mice(
    data = dat[,c(14, 16)],
    m = 5,
    maxit = 50,
    method = 'pmm',
    seed = 1234,
    printFlag = FALSE
  )

set.seed(1234)  # set seed for random imputed dataset selection
compdat <- complete(impdat, round(runif(1, 1, 5), 0))
# randomly select imputed dataset and combine with complete dataset

# inspect imputed data
# imputed datasets in magenta; observed in blue. are values plausible?
# stripplot(impdat)
# do imputed points fall in likely area of distrib.?

dat[,c(14, 16)] <- compdat[,1:ncol(compdat)]

# ..Nitrogen ------------------------------------

impdat <-
  mice(
    data = dat[,c(11:13, 15, 17)],
    m = 5,
    maxit = 50,
    method = 'pmm',
    seed = 1234,
    printFlag = FALSE
  )

set.seed(1234)  # set seed for random imputed dataset selection
compdat <- complete(impdat, round(runif(1, 1, 5), 0))
# randomly select imputed dataset and combine with complete dataset

# inspect imputed data
# imputed datasets in magenta; observed in blue. are values plausible?
#stripplot(impdat)
# do imputed points fall in likely area of distrib.?

dat[,c(11:13, 15, 17)] <- compdat[,1:ncol(compdat)]

# ..No3 -----------------------------------

impdat <-
  mice(
    data = dat[,c(11:13, 15, 19)],
    m = 5,
    maxit = 50,
    method = 'pmm',
    seed = 1234,
    printFlag = FALSE
  )

set.seed(1234)  # set seed for random imputed dataset selection
compdat <- complete(impdat, round(runif(1, 1, 5), 0))
# randomly select imputed dataset and combine with complete dataset

# inspect imputed data
# imputed datasets in magenta; observed in blue. are values plausible?
#stripplot(impdat)
# do imputed points fall in likely area of distrib.?

dat[,c(11:13, 15, 19)] <- compdat[,1:ncol(compdat)]

rm(impdat)
rm(compdat)

# repair MDLs -------------------------------------------------------------
if (sum(dat$TSS.mgl < 5, na.rm = TRUE) / sum(!is.na(dat$TSS.mgl)) < 0.05) {
  print("Less than 5% of TSS values are below MDL")
  print(sum(dat$TSS.mgl < 5, na.rm = TRUE) / sum(!is.na(dat$TSS.mgl)) *100)
  dat <- dat %>% mutate(TSS.mgl = replace(TSS.mgl, TSS.mgl < 5, 5/2))
} else {
  print("More than 5% of TSS values are below MDL and")
  print("the values have remained unchanged")
  print(dat %>% count(TSS.mgl < 5))
  print(sum(dat$TSS.mgl < 5, na.rm = TRUE) / sum(!is.na(dat$TSS.mgl)) *100)}


if (sum(dat$TUR.ntu < .02, na.rm = TRUE) / sum(!is.na(dat$TUR.ntu)) < 0.05) {
  print("Less than 5% of TUR values are below MDL")
  print(sum(dat$TUR.ntu < .02, na.rm = TRUE) / sum(!is.na(dat$TUR.ntu)) *100)
  dat <- dat %>% mutate(TUR.ntu = replace(TUR.ntu, TUR.ntu < .02, .02/2))
} else {
  print("More than 5% of TUR values are below MDL and")
  print("the values have remained unchanged")
  print(dat %>% count(TUR.ntu < .02))
  print(sum(dat$TUR.ntu < .02, na.rm = TRUE) / sum(!is.na(dat$TUR.ntu)) *100)}


if (sum(dat$NO3.NO2.mgl < .23, na.rm = TRUE) / sum(!is.na(dat$NO3.NO2.mgl)) < 0.05) {
  print("Less than 5% of NO3.NO2 values are below MDL")
  print(sum(dat$NO3.NO2.mgl < .23, na.rm = TRUE) / sum(!is.na(dat$NO3.NO2.mgl)) *100)
  dat <- dat %>% mutate(NO3.NO2.mgl = replace(NO3.NO2.mgl, NO3.NO2.mgl < .23, .23/2))
} else {
  print("More than 5% of NO3.NO2 values are below MDL and")
  print("the values have remained unchanged")
  print(dat %>% count(NO3.NO2.mgl < .23))
  print(sum(dat$NO3.NO2.mgl < .23, na.rm = TRUE) / sum(!is.na(dat$NO3.NO2.mgl)) *100)}


if (sum(dat$TKN.mgl < .77, na.rm = TRUE) / sum(!is.na(dat$TKN.mgl)) < 0.05) {
  print("Less than 5% of TKN values are below MDL")
  print(sum(dat$TKN.mgl < .77, na.rm = TRUE) / sum(!is.na(dat$TKN.mgl)) *100)
  dat <- dat %>% mutate(TKN.mgl = replace(TKN.mgl, TKN.mgl < .77, .77/2))
} else {
  print("More than 5% of TKN values are below MDL and")
  print("the values have remained unchanged")
  print(dat %>% count(TKN.mgl < .77))
  print(sum(dat$TKN.mgl < .77, na.rm = TRUE) / sum(!is.na(dat$TKN.mgl)) * 100)
  }


if (sum(dat$TN.mgl < 1, na.rm = TRUE) / sum(!is.na(dat$TN.mgl)) < 0.05) {
  print("Less than 5% of TN values are below MDL")
  print(sum(dat$TN.mgl < 1, na.rm = TRUE) / sum(!is.na(dat$TN.mgl)) *100)
  dat <- dat %>% mutate(TN.mgl = replace(TN.mgl, TN.mgl < 1, 1/2))
} else {
  print("More than 5% of TN values are below MDL and")
  print("the values have remained unchanged")
  print(dat %>% count(TN.mgl < 1))
  print(sum(dat$TN.mgl < 1, na.rm = TRUE) / sum(!is.na(dat$TN.mgl)) *100)
  }


if (sum(dat$TIP.mgl < .05, na.rm = TRUE) / sum(!is.na(dat$TIP.mgl)) < 0.05) {
  print("Less than 5% of TIP values are below MDL")
  print(sum(dat$TIP.mgl < .05, na.rm = TRUE) / sum(!is.na(dat$TIP.mgl)) *100)
  dat <- dat %>% mutate(TIP.mgl = replace(TIP.mgl, TIP.mgl < .05, .05/2))
} else {
  print("More than 5% of TIP values are below MDL and")
  print("the values have remained unchanged")
  print(dat %>% count(TIP.mgl < .05))
  print(sum(dat$TIP.mgl < .05, na.rm = TRUE) / sum(!is.na(dat$TIP.mgl)) *100)}


if (sum(dat$NH3.mgl < .01, na.rm = TRUE) / sum(!is.na(dat$NH3.mgl)) < 0.05) {
  print("Less than 5% of NH3 values are below MDL")
  print(sum(dat$NH3.mgl < .01, na.rm = TRUE) / sum(!is.na(dat$NH3.mgl)) *100)
  dat <- dat %>% mutate(NH3.mgl = replace(NH3.mgl, NH3.mgl < .01, .01/2))
} else {
  print("More than 5% of NH3 values are below MDL and")
  print("the values have remained unchanged")
  print(dat %>% count(NH3.mgl < .01))
  print(sum(dat$NH3.mgl < .01, na.rm = TRUE) / sum(!is.na(dat$NH3.mgl)) *100)}


if (sum(dat$OrthoP.mgl < .01, na.rm = TRUE) / sum(!is.na(dat$OrthoP.mgl)) < 0.05) {
  print("Less than 5% of Ortho Phosphorus values are below MDL")
  print(sum(dat$OrthoP.mgl < .01, na.rm = TRUE) / sum(!is.na(dat$OrthoP.mgl)) *100)
  dat <- dat %>% mutate(OrthoP.mgl = replace(OrthoP.mgl, OrthoP.mgl < .01, .01/2))
} else {
  print("More than 5% of Ortho Phosphorus values are below MDL and")
  print("the values have remained unchanged")
  print(dat %>% count(OrthoP.mgl < .01))
  print(sum(dat$OrthoP.mgl < .01, na.rm = TRUE) / sum(!is.na(dat$OrthoP.mgl)) *100)}


if (sum(dat$NOx.mgl < .007, na.rm = TRUE) / sum(!is.na(dat$NOx.mgl)) < 0.05) {
  print("Less than 5% of NOx values are below MDL")
  print(sum(dat$NOx.mgl < .007, na.rm = TRUE) / sum(!is.na(dat$NOx.mgl)) *100)
  dat <- dat %>% mutate(NOx.mgl = replace(NOx.mgl, NOx.mgl < .007, .007/2))
} else {
  print("More than 5% of NOx values are below MDL and")
  print("the values have remained unchanged")
  print(dat %>% count(NOx.mgl < .007))
  print(sum(dat$NOx.mgl < .007, na.rm = TRUE) / sum(!is.na(dat$NOx.mgl)) *100)}


if (sum(dat$NO2.mgl < .05, na.rm = TRUE) / sum(!is.na(dat$NO2.mgl)) < 0.05) {
  print("Less than 5% of NO2 values are below MDL")
  print(sum(dat$NO2.mgl < .05, na.rm = TRUE) / sum(!is.na(dat$NO2.mgl)) *100)
  dat <- dat %>% mutate(NO2.mgl = replace(NO2.mgl, NO2.mgl < .05, .05/2))
} else {
  print("More than 5% of NO2 values are below MDL and")
  print("the values have remained unchanged")
  print(dat %>% count(NO2.mgl < .05))
  print(sum(dat$NO2.mgl < .05, na.rm = TRUE) / sum(!is.na(dat$NO2.mgl)) *100)}


if (sum(dat$NO3.mgl < .0007, na.rm = TRUE) / sum(!is.na(dat$NO3.mgl)) < 0.05) {
  print("Less than 5% of NO3 values are below MDL")
  print(sum(dat$NO3.mgl < .0007, na.rm = TRUE) / sum(!is.na(dat$NO3.mgl)) *100)
  dat <- dat %>% mutate(NO3.mgl = replace(NO3.mgl, NO3.mgl < .0007, .0007/2))
} else {
  print("More than 5% of NO3 values are below MDL and")
  print("the values have remained unchanged")
  print(dat %>% count(NO3.mgl < .0007))
  print(sum(dat$NO3.mgl < .0007, na.rm = TRUE) / sum(!is.na(dat$NO3.mgl)) *100)}

#another round of imput for ammonia and tkn
#NH3 ---------------

dat$NH3.mgl <-
  ifelse(
    dat$NH3.mgl < .01,
    NA,
    dat$NH3.mgl
  )

impdat <-
  mice(
    data = dat[,c(11:13, 15, 17)],
    m = 5,
    maxit = 50,
    method = 'pmm',
    seed = 1234,
    printFlag = FALSE
  )

set.seed(1234)  # set seed for random imputed dataset selection
compdat <- complete(impdat, round(runif(1, 1, 5), 0))
# randomly select imputed dataset and combine with complete dataset

# inspect imputed data
# imputed datasets in magenta; observed in blue. are values plausible?
#stripplot(impdat)
# do imputed points fall in likely area of distrib.?

dat[,c(11:13, 15, 17)] <- compdat[,1:ncol(compdat)]

# TKN-------------------
print(
  paste(
    sum(dat$TKN.mgl < 0, na.rm = TRUE) / sum(!is.na(dat$TKN.mgl)) *100,
    "% of TKN values are below 0"
  )
)


dat$TKN.mgl <-
  ifelse(
    dat$TKN.mgl < 0,
    NA,
    dat$TKN.mgl
  )

summary(dat$TKN.mgl)

impdat <-
  mice(
    data = dat[,c(11:13, 15, 17)],
    m = 5,
    maxit = 50,
    method = 'pmm',
    seed = 1234,
    printFlag = FALSE
  )

set.seed(1234)  # set seed for random imputed dataset selection
compdat <- complete(impdat, round(runif(1, 1, 5), 0))
# randomly select imputed dataset and combine with complete dataset

# inspect imputed data
# imputed datasets in magenta; observed in blue. are values plausible?
#stripplot(impdat)
# do imputed points fall in likely area of distrib.?

dat[,c(11:13, 15, 17)] <- compdat[,1:ncol(compdat)]



```

```{r, echo= FALSE, message= FALSE, warning= FALSE}
INST <- dat[dat$sample.type== "INST",]%>% 
  rename(LOCCODE = sampleid,
         TSS= TSS.mgl,
         TUR = TUR.ntu,
         NN = NO3.NO2.mgl,
         TKN= TKN.mgl,
         TN = TN.mgl,
         TIP = TIP.mgl,
         NH3 = NH3.mgl,
         DIP = OrthoP.mgl
         )

oldINST <- read.csv("combinedgrabs.csv") %>% 
 rename(Date = 1,
        TUR = 13)  %>% 
  mutate(
         Date = lubridate::mdy(Date),
         Alk = as.numeric(Alk),
         NH3 = as.numeric(NH3),
         TSS = as.numeric(TSS),
         NN = as.numeric(NN),
         TP = as.numeric(TP),
         TN = as.numeric(TN),
         OrgN = as.numeric(OrgN),
         TOC = as.numeric(TOC),
         COD = as.numeric(COD),
         Cl = as.numeric(Cl),
         TKN = as.numeric(TKN)
         )
  
instreamdata <- full_join(INST, oldINST)
instreamdata <- instreamdata[,-c(4:8, 20:31, 39:41)]
instreamdata$LOCCODE <- as.factor(instreamdata$LOCCODE)


attach(instreamdata)
oldharris <- instreamdata[LOCCODE== "HN3"|LOCCODE== "HN4"|LOCCODE== "HN5"
                 |LOCCODE== "HN6"|LOCCODE== "HNO"|LOCCODE== "HNSTA"
                 |LOCCODE== "HNT"|LOCCODE== "HS1"|LOCCODE== "HS2",]

oldporter <- instreamdata[LOCCODE== "PN1"|LOCCODE== "PN2"|LOCCODE== "PS1"|
                  LOCCODE== "PS2"|LOCCODE== "PS3"|LOCCODE== "PS4"|
                  LOCCODE== "PS5"|LOCCODE== "PS6",]
detach(instreamdata)

```

\newpage

# TURBIDITY

## Harris Bayou

### Historic Harris Bayou sites

```{r, warning= FALSE, echo=FALSE}
a <- summarise(oldharris, 
               n = count(!is.na(oldharris$TUR)),
               Mean= mean(TUR, na.rm=TRUE),
               Median= median(TUR, na.rm = TRUE))
 
a %>%  kable()

df1 <- data.frame("n"= NA, "Mean"= NA, "Median"= NA)
df1 <- rbind(df1, a)
```

### Data from Harris Bayou site associated with Delta FARM project

```{r, echo=FALSE}
b <- instreamdata %>% 
  filter(LOCCODE =="HB") %>% 
summarise(n= count(!is.na(TUR)),
          Mean= mean(TUR, na.rm= TRUE),
          Median= median(TUR, na.rm = TRUE))
  
b %>% kable()
df1 <- rbind(df1, b)
```

### Plot the data points. 

Red indicates our recent data.

```{r fig1, results='hold', warning=FALSE, echo= FALSE, fig.height= 3 }
ggplot()+
  geom_jitter(data= oldharris, aes(y= TUR, x= LOCCODE))+
  geom_jitter(data= instreamdata[instreamdata$LOCCODE == "HB",], aes(y= TUR, x= LOCCODE, color= "red"))+
  theme(legend.position = "none")+
  labs(x ="Location", y = "Turbidity (ntu)")
  
  
```

```{r fig2, warning=FALSE, fig.height= 3, echo= FALSE}
a <- ggplot()+
  geom_boxplot(data= oldharris, aes(y= TUR, x= LOCCODE))+
  geom_boxplot(data= instreamdata[instreamdata$LOCCODE == "HB",], aes(y= TUR, x= LOCCODE, color= "red"))+
  theme(legend.position = "none", axis.text.x = element_text(angle = 45))+
  labs(x ="Location", y = "Turbidity (ntu)")
b <- ggplot()+
  geom_boxplot(data= instreamdata[instreamdata$LOCCODE == "HB",], aes(y= TUR, x= LOCCODE, color= "red"))+
  geom_boxplot(data= oldharris, aes(y= TUR, x= "Historic combined"))+
  theme(legend.position = "none", axis.title.y = element_blank())+
  labs(x ="Location", y = "Turbidity (ntu)")

a+b
```

\newpage

### Test for a difference in means

```{r}
wilcox.test(y= oldharris$TUR, x= instreamdata[instreamdata$LOCCODE == "HB", "TUR"],
            alternative = "less")

```

P-value < $\alpha$ here indicates that we can reject the null hypothesis that the difference is not less than zero and conclude, at a %95 confidence level, that the values of our recent data is in fact lower than the Historic data.  $\alpha = 0.05$ is used throughout.



### Percent reduction in the median value
```{r}
x <- median(oldharris$TUR, na.rm= TRUE)
y <- median(instreamdata[instreamdata$LOCCODE == "HB", "TUR"])
z <- (y-x)/x * 100
z
```
 
Here we would use median values because there are outliers and apparent skewness from the boxplot. The mean reduction is even greater (%72). Negative value represents a decrease relative to the Historic data.

```{r, warning= FALSE, echo=FALSE}
df <- data.frame("Method" = NA, "Analyte"= NA,"Location" = NA, "Alternative" = NA, "P.value" = NA, "relative.reduction"= NA)
df$Location <- "Harris Bayou"
df$Analyte <- "Turbidity"
df$Method <- "Wilcox rank sum test"
df$Alternative <- wilcox.test(y= oldharris$TUR, x= instreamdata[instreamdata$LOCCODE == "HB", "TUR"],
            alternative = "less")$alternative
df$P.value <- wilcox.test(y= oldharris$TUR, x= instreamdata[instreamdata$LOCCODE == "HB", "TUR"],
            alternative = "less")$p.value
df$relative.reduction <- z
```

\newpage

## Porter Bayou

### Historic Porter Bayou sites

```{r, warning= FALSE, echo=FALSE}
c <- summarise(oldporter,
               n = count(!is.na(oldporter$TUR)),
               Mean= mean(TUR, na.rm=TRUE),
               Median= median(TUR, na.rm = TRUE))
c %>% kable()
df1 <- rbind(df1, c)
```

### Data from Porter Bayou site associated with Delta FARM project

```{r, echo=FALSE}
d <- instreamdata %>% 
  filter(LOCCODE =="PB") %>% 
summarise(n= count(!is.na(TUR)),
          Mean= mean(TUR, na.rm= TRUE), 
          Median= median(TUR, na.rm = TRUE))
d %>% kable()
df1 <- rbind(df1, d)
```

### Plot the data points.

Red indicates our recent data.

```{r fig3, results='hold', warning=FALSE, echo= FALSE, fig.height= 3}
ggplot()+
  geom_jitter(data= oldporter, aes(y= TUR, x= LOCCODE))+
  geom_jitter(data= instreamdata[instreamdata$LOCCODE == "PB",], aes(y= TUR, x= LOCCODE, color= "red"))+
  theme(legend.position = "none")+
  labs(x ="Location", y = "Turbidity (ntu)")
  
```

```{r fig4, warning=FALSE, fig.height=3, echo= FALSE}
a <- ggplot()+
  geom_boxplot(data= oldporter, aes(y= TUR, x= LOCCODE))+
  geom_boxplot(data= instreamdata[instreamdata$LOCCODE == "PB",], aes(y= TUR, x= LOCCODE, color= "red"))+
  theme(legend.position = "none", axis.text.x = element_text(angle = 45))+
  labs(x ="Location", y = "Turbidity (ntu)")
b <- ggplot()+
  geom_boxplot(data= instreamdata[instreamdata$LOCCODE == "PB",], aes(y= TUR, x= LOCCODE, color= "red"))+
  geom_boxplot(data= oldporter, aes(y= TUR, x= "Historic combined"))+
  theme(legend.position = "none", axis.title.y = element_blank())+
  labs(x ="Location", y = "Turbidity (ntu)")

a+b
```

\newpage

### Test for a difference in means

```{r}
wilcox.test(y= oldporter$TUR, x= instreamdata[instreamdata$LOCCODE == "PB", "TUR"],
            alternative = "greater")
```

P-value < $\alpha$ here indicates that we can reject the null hypothesis that the difference is not greater than zero and conclude, at a %95 confidence level, that the values of our recent data is higher than the historic data.  
$H_0:$ old data $\geq$ new data  
$H_a:$ old data $<$ new data

### Percent reduction in the median value

```{r}
x <- median(oldporter$TUR, na.rm= TRUE)
y <- median(instreamdata[instreamdata$LOCCODE == "PB", "TUR"])
z <- (y-x)/x * 100
z
```
 
Here we would use median values because there are outliers and apparent skewness from the boxplot. Positive values represents a relative increase. This relative change can range from (-) 100% decrease to (+) infinite increase.

```{r, warning= FALSE, echo=FALSE}
df <- rbind(df, list("Wilcox rank sum test",
              "Turbidity",
              "Porter Bayou", 
            wilcox.test(y= oldporter$TUR, x= instreamdata[instreamdata$LOCCODE == "PB", "TUR"],
              alternative = "greater")$alternative,
            wilcox.test(y= oldporter$TUR, x= instreamdata[instreamdata$LOCCODE == "PB", "TUR"],
              alternative = "greater")$p.value, z
            )
      )

```

\newpage

# TOTAL SUSPENDED SOLIDS

## Harris Bayou

### Historic Harris Bayou sites

```{r, warning= FALSE, echo=FALSE}
e <- summarise(oldharris,
               n = count(!is.na(oldharris$TSS)),
               Mean= mean(TSS, na.rm=TRUE),
               Median= median(TSS, na.rm = TRUE))
e %>% kable()
df1 <- rbind(df1, e)
```

### Data from Harris Bayou site associated with Delta FARM project

```{r, echo=FALSE}
f <- instreamdata %>% 
  filter(LOCCODE =="HB") %>% 
summarise(n= count(!is.na(TSS)),
          Mean= mean(TSS, na.rm= TRUE),
          Median= median(TSS, na.rm = TRUE))

f %>% kable()
df1 <- rbind(df1, f)
```

### Plot the data points. 

Red indicates our recent data.

```{r fig5, results='hold', warning=FALSE, echo= FALSE, fig.height= 3 }
ggplot()+
  geom_jitter(data= oldharris, aes(y= TSS, x= LOCCODE))+
  geom_jitter(data= instreamdata[instreamdata$LOCCODE == "HB",], aes(y= TSS, x= LOCCODE, color= "red"))+
  theme(legend.position = "none")+
  labs(x ="Location", y = "TSS (mg/L)")
  
```

```{r fig6, warning=FALSE, fig.height= 3, echo= FALSE}
a <- ggplot()+
  geom_boxplot(data= oldharris, aes(y= TSS, x= LOCCODE))+
  geom_boxplot(data= instreamdata[instreamdata$LOCCODE == "HB",], aes(y= TSS, x= LOCCODE, color= "red"))+
  theme(legend.position = "none", axis.text.x = element_text(angle = 45))+
  labs(x ="Location", y = "TSS (mg/L)")
b <- ggplot()+
  geom_boxplot(data= instreamdata[instreamdata$LOCCODE == "HB",], aes(y= TSS, x= LOCCODE, color= "red"))+
  geom_boxplot(data= oldharris, aes(y= TSS, x= "Historic combined"))+
  theme(legend.position = "none", axis.title.y = element_blank())+
  labs(x ="Location", y = "TSS (mg/L)")

a+b
```

\newpage

### Test for a difference in means

```{r}
wilcox.test(y= oldharris$TSS, x= instreamdata[instreamdata$LOCCODE == "HB", "TSS"],
            alternative = "less")
```

P-value < $\alpha = 0.05$ means we can conclude the alternative hypothesis. 
$H_a:$ our recent samples had less total suspended solids than our historic observations

### Percent reduction in the median value
```{r}
x <- median(oldharris$TSS, na.rm= TRUE)
y <- median(instreamdata[instreamdata$LOCCODE == "HB", "TSS"])
z <- (y-x)/x * 100
z
```
```{r, warning= FALSE, echo=FALSE}
df <- rbind(df, list("Wilcox rank sum test",
              "Total Suspended Solids",
              "Harris Bayou", 
            wilcox.test(y= oldharris$TSS, x= instreamdata[instreamdata$LOCCODE == "HB", "TSS"],
              alternative = "less")$alternative,
            wilcox.test(y= oldharris$TSS, x= instreamdata[instreamdata$LOCCODE == "HB", "TSS"],
              alternative = "less")$p.value, z
            )
      )

```
 
Negative value represents a reduction relative to the historic observations.

\newpage

## Porter Bayou

### Historic Porter Bayou sites

```{r, warning= FALSE, echo=FALSE}
g <- summarise(oldporter,
               n = count(!is.na(oldporter$TSS)),
               Mean= mean(TSS, na.rm=TRUE),
               Median= median(TSS, na.rm = TRUE))
g %>% kable()
df1 <- rbind(df1, g)
```

### Data from Porter Bayou site associated with the original Delta FARM project

```{r, echo=FALSE}
h <- instreamdata %>% 
  filter(LOCCODE =="PB") %>% 
summarise(n= count(!is.na(TSS)),
          Mean= mean(TSS, na.rm= TRUE), 
          Median= median(TSS, na.rm = TRUE))
h %>% kable()
df1 <- rbind(df1, h)
```

### Plot the data points. 

Red indicates our recent data.

```{r fig7, results='hold', warning=FALSE, echo= FALSE, fig.height= 3}
ggplot()+
  geom_jitter(data= oldporter, aes(y= TSS, x= LOCCODE))+
  geom_jitter(data= instreamdata[instreamdata$LOCCODE == "PB",], aes(y= TSS, x= LOCCODE, color= "red"))+
  theme(legend.position = "none")+
  coord_cartesian(ylim = c(0,2500))+
  labs(x ="Location", y = "TSS (mg/L)")
```

```{r fig8, warning=FALSE, fig.height=3, echo= FALSE}
a <- ggplot()+
  geom_boxplot(data= oldporter, aes(y= TSS, x= LOCCODE))+
  geom_boxplot(data= instreamdata[instreamdata$LOCCODE == "PB",], aes(y= TSS, x= LOCCODE, color= "red"))+
  theme(legend.position = "none", axis.text.x = element_text(angle = 45))+
  coord_cartesian(ylim = c(0,2500))+
  labs(x ="Location", y = "TSS (mg/L)")

b <- ggplot()+
  geom_boxplot(data= instreamdata[instreamdata$LOCCODE == "PB",], aes(y= TSS, x= LOCCODE, color= "red"))+
  geom_boxplot(data= oldporter, aes(y= TSS, x= "Historic combined"))+
  theme(legend.position = "none", axis.title.y = element_blank())+
  coord_cartesian(ylim = c(0,2500))+
  labs(x ="Location", y = "TSS (mg/L)")


a+b
```

  $*$ some outliers clipped from visualization
  
\newpage

### Test for a difference in means

```{r}
wilcox.test(y= oldporter$TSS, x= instreamdata[instreamdata$LOCCODE == "PB", "TSS"],
            alternative = "greater")
```
P-value < $\alpha$ leads to rejection of the null hypothesis.  $H_a:$ new data is greater than the old data
The test lends statistical significance to old porter bayou data values being less than what we recently collected.

### Percent reduction in the median value

```{r}
x <- median(oldporter$TSS, na.rm= TRUE)
y <- median(instreamdata[instreamdata$LOCCODE == "PB", "TSS"])
z <- (y-x)/x * 100
z
```
 
This positive number represents a four fold increase from values formerly observed in porter bayou.

```{r, warning= FALSE, echo=FALSE}
df <- rbind(df, list("Wilcox rank sum test",
              "Total Suspended Solids",
              "Porter Bayou", 
            wilcox.test(y= oldporter$TSS, x= instreamdata[instreamdata$LOCCODE == "PB", "TSS"],
              alternative = "greater")$alternative,
            wilcox.test(y= oldporter$TSS, x= instreamdata[instreamdata$LOCCODE == "PB", "TSS"],
              alternative = "greater")$p.value, z
            )
      )

```
 
\newpage

# TOTAL NITROGEN

## Harris Bayou

### Historic Harris Bayou sites

```{r, warning= FALSE, echo=FALSE}
i <- summarise(oldharris,
               n = count(!is.na(oldharris$TN)),
               Mean= mean(TN, na.rm=TRUE),
               Median= median(TN, na.rm = TRUE))
i %>% kable()
df1 <- rbind(df1, i)
```

### Data from Harris Bayou site associated with Delta FARM project

```{r, echo=FALSE}
j <- instreamdata %>% 
  filter(LOCCODE =="HB") %>% 
summarise(n= count(!is.na(TN)),
          Mean= mean(TN, na.rm= TRUE),
          Median= median(TN, na.rm = TRUE))
j %>% kable()
df1 <- rbind(df1, j)
```

### Plot the data points. 

Red indicates our recent data.

```{r fig9, results='hold', warning=FALSE, echo= FALSE, fig.height= 3 }
ggplot()+
  geom_jitter(data= oldharris, aes(y= TN, x= LOCCODE))+
  geom_jitter(data= instreamdata[instreamdata$LOCCODE == "HB",], aes(y= TN, x= LOCCODE, color= "red"))+
  theme(legend.position = "none")+
  labs(x ="Location", y = "TN (mg/L)")
  
```

```{r fig10, warning=FALSE, fig.height= 3, echo= FALSE}
a <- ggplot()+
  geom_boxplot(data= oldharris, aes(y= TN, x= LOCCODE))+
  geom_boxplot(data= instreamdata[instreamdata$LOCCODE == "HB",], aes(y= TN, x= LOCCODE, color= "red"))+
  theme(legend.position = "none", axis.text.x = element_text(angle = 45))+
  labs(x ="Location", y = "TN (mg/L)")
b <- ggplot()+
  geom_boxplot(data= instreamdata[instreamdata$LOCCODE == "HB",], aes(y= TN, x= LOCCODE, color= "red"))+
  geom_boxplot(data= oldharris, aes(y= TN, x= "Historic combined"))+
  theme(legend.position = "none", axis.title.y = element_blank())+
  labs(x ="Location", y = "TN (mg/L)")

a+b
```

\newpage

### Test for a difference in means

```{r}
wilcox.test(y= oldharris$TN, x= instreamdata[instreamdata$LOCCODE == "HB", "TN"],
            alternative = "less")
```

P-value  $< \alpha =0.05:$ reject null that old values are $\leq$ new observations.  
$H_a:$ new observations < old observations.

### Percent reduction in the median value
```{r}
x <- median(oldharris$TN, na.rm= TRUE)
y <- median(instreamdata[instreamdata$LOCCODE == "HB", "TN"])
z <- (y-x)/x * 100
z
```

 Negative value here represents a twenty percent reduction over the historic observations.

```{r, warning= FALSE, echo=FALSE}
df <- rbind(df, list("Wilcox rank sum test",
              "Total Nitrogen",
              "Harris Bayou", 
            wilcox.test(y= oldharris$TN, x= instreamdata[instreamdata$LOCCODE == "HB", "TN"],
              alternative = "less")$alternative,
            wilcox.test(y= oldharris$TN, x= instreamdata[instreamdata$LOCCODE == "HB", "TN"],
              alternative = "less")$p.value, z
            )
      )

```

\newpage

## Porter Bayou

### Historic Porter Bayou sites

```{r, warning= FALSE, echo=FALSE}
k <- summarise(oldporter,
               n = count(!is.na(oldporter$TN)),
               Mean= mean(TN, na.rm=TRUE),
               Median= median(TN, na.rm = TRUE))
k %>% kable()
df1 <- rbind(df1, k)
```

### Data from Porter Bayou site associated with Delta FARM project

```{r, echo=FALSE}
l <- instreamdata %>% 
  filter(LOCCODE =="PB") %>% 
summarise(n= count(!is.na(TN)),
          Mean= mean(TN, na.rm= TRUE), 
          Median= median(TN, na.rm = TRUE))
l %>% kable()
df1 <- rbind(df1, l)
```

### Plot the data points.

Red indicates our recent data.

```{r fig11, results='hold', warning=FALSE, echo= FALSE, fig.height= 3}
ggplot()+
  geom_jitter(data= oldporter, aes(y= TN, x= LOCCODE))+
  geom_jitter(data= instreamdata[instreamdata$LOCCODE == "PB",], aes(y= TN, x= LOCCODE, color= "red"))+
  theme(legend.position = "none")+
  labs(x ="Location", y = "TN (mg/L)")
  
#coord_cartesian(ylim = c(0,2500))
```

```{r fig12, warning=FALSE, fig.height=3, echo= FALSE}
a <- ggplot()+
  geom_boxplot(data= oldporter, aes(y= TN, x= LOCCODE))+
  geom_boxplot(data= instreamdata[instreamdata$LOCCODE == "PB",], aes(y= TN, x= LOCCODE, color= "red"))+
  theme(legend.position = "none", axis.text.x = element_text(angle = 45))+
  labs(x ="Location", y = "TN (mg/L)")
    #coord_cartesian(ylim = c(0,2500))

b <- ggplot()+
  geom_boxplot(data= instreamdata[instreamdata$LOCCODE == "PB",], aes(y= TN, x= LOCCODE, color= "red"))+
  geom_boxplot(data= oldporter, aes(y= TN, x= "Historic combined"))+
  theme(legend.position = "none", axis.title.y = element_blank())+
  labs(x ="Location", y = "TN (mg/L)")
  #coord_cartesian(ylim = c(0,2500))


a+b
```

\newpage

### Test for a difference in means

```{r}
wilcox.test(y= oldporter$TN, x= instreamdata[instreamdata$LOCCODE == "PB", "TN"],
            alternative = "greater")
```

P-value < $\alpha$ leads to rejection of null hypothesis that porter bayou TN concentrations are $\leq$ old observations and conclude the alternative. <b> $H_a:$ new observations are greater than old ones.

### Percent reduction in the median value

```{r}
x <- median(oldporter$TN, na.rm= TRUE)
y <- median(instreamdata[instreamdata$LOCCODE == "PB", "TN"])
z <- (y-x)/x * 100
z
```
 
 This positive value represents an increase from our historic observations.

```{r, warning= FALSE, echo=FALSE}
df <- rbind(df, list("Wilcox rank sum test",
              "Total Nitrogen",
              "Porter Bayou", 
            wilcox.test(y= oldporter$TN, x= instreamdata[instreamdata$LOCCODE == "PB", "TN"],
              alternative = "greater")$alternative,
            wilcox.test(y= oldporter$TN, x= instreamdata[instreamdata$LOCCODE == "PB", "TN"],
              alternative = "greater")$p.value, z
            )
      )

```
 
\newpage

# Summary

There is a statistically significant ($\alpha = 0.05$) departure for all of our analyte location combinations. Though the direction of this difference was opposite in our two watersheds. Historic data in Porter Bayou contained lower sediment and nutrient concentrations while our current samples in Harris Bayou had lower concentrations than the historic observations.  
<b>  
Perhaps here we could go into a brief description of the sample locations and the BMPs used historically and currently.


```{r, results= 'asis', echo= FALSE}
df1 <- df1[-1,]
df1$Analyte <- "Turbidity" 
df1$Analyte[5:8] <- "Total Suspended Solids"
df1$Analyte[9:12] <- "Total Nitrogen"
df1$Location <- c("Harris Bayou", "Harris Bayou", "Porter Bayou", "Porter Bayou",
                  "Harris Bayou", "Harris Bayou", "Porter Bayou", "Porter Bayou",
                  "Harris Bayou", "Harris Bayou", "Porter Bayou", "Porter Bayou")
df1$Observation <- c("Historic", "Current", "Historic", "Current",
                     "Historic", "Current", "Historic", "Current",
                     "Historic", "Current", "Historic", "Current")
df1 <- df1[,c(4,5,6,1,2,3)]
row.names(df1) <- NULL
df1 %>% kable(format = "latex", booktabs= TRUE, digits = 2, caption = "Summary of sample concentration means") %>% 
  collapse_rows(columns = c(1,2)) %>%
  kable_styling(latex_options =  "hold_position")

```


```{r, results= 'asis', echo= FALSE}
df <- df[,c(1,3,2,4,5,6)]
df %>% 
  arrange(Location) %>% 
  kable(format = "latex", booktabs= TRUE, caption = "Comparison of current observations with historic data") %>% 
  collapse_rows(columns = c(1,2)) %>%
  kable_styling(latex_options =  "hold_position")

```


Alternative specifies $H_a$: current observations are ___ than historical observations. Positive relative reduction values represent an increase over historical observations. Relative.reduction can range between 100% reduction and $\infty$ increase.

